---
title: "ALCL_subtype_assignment"
author: "Putri Ramadani"
date: "2026-01-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Load necessary library
```{r}
library(data.table)
library(readxl)
library(dplyr)
library(stringr)
library(ggplot2)
library(ComplexHeatmap)
library(circlize)
library(grid)

# edgeR & singscore
if (!requireNamespace("edgeR", quietly = TRUE)) install.packages("edgeR")
library(edgeR)
devtools::install_github('DavisLaboratory/singscore')
library(singscore)

# Optional: read GEO SOFT locally
BiocManager::install("GEOquery", update = FALSE, ask = FALSE)
library(GEOquery)

```

2. Paths and directory
```{r}
getwd()
# [1] "/Users/putriramadani/Documents/GitHub/ALCL_subtyping/script"
root <- normalizePath("..")
root
# [1] "/Users/putriramadani/Documents/GitHub/ALCL_subtyping"

data_dir <- file.path(root, "data")
fig_dir  <- file.path(root, "figures")
res_dir  <- file.path(root, "result")

counts_file <- file.path(data_dir, "GSE280480_20241017-RawGeneCounts-For-GEO.txt")
soft_file   <- file.path(data_dir, "GSE280480_family.soft")
gene_xlsx   <- file.path(data_dir, "ALCL_subtype_gene_list.xlsx")
```

3. Load raw counts
```{r}
counts_dt <- read.delim(counts_file, header = FALSE, check.names = FALSE)

new_header <- as.character(unlist(counts_dt[1, ]))
counts_dt2 <- counts_dt[-1, ]
colnames(counts_dt2) <- new_header

genes <- counts_dt2[[1]]
counts <- as.matrix(counts_dt2[, -1])
rownames(counts) <- genes
mode(counts) <- "numeric"

# check:
dim(counts) 
# [1] 64253   393
head(colnames(counts), 10)   
# [1] "53"  "59"  "88"  "91"  "133" "178" "190" "237" "238" "240"
counts[1:3, 1:4]
#                                        53  59  88  91
# ENSG00000223972@DDX11L1@11869@14409     0   0   2   1
# ENSG00000227232@WASH7P@14404@29570    350 386 622 530
# ENSG00000278267@MIR6859-1@17369@17436   0  59   0   0
summary(colSums(counts))    
#     Min.   1st Qu.    Median      Mean   3rd Qu.      Max. 
#  4549550  33782211  47561129  50734564  62759455 145752710 


```

4. Convert gene IDs to gene symbols (and sum duplicates) to match gene list on excel file
```{r}
# WHAT: Convert "ENSG@SYMBOL@start@end" to SYMBOL and collapse duplicate symbols by SUM
# WHY: My signature gene list is in gene symbols; summing keeps all signal

genes_raw <- counts_dt2[[1]]
symbol <- sub("^[^@]+@([^@]+)@.*$", "\\1", genes_raw)

counts_num <- counts_dt2[, -1, drop = FALSE]
counts_num[] <- lapply(counts_num, as.numeric)

counts_sym <- aggregate(. ~ symbol,
                        data = cbind(symbol = symbol, counts_num),
                        FUN = sum)

rownames(counts_sym) <- counts_sym$symbol
counts <- as.matrix(counts_sym[, -1, drop = FALSE])

# Note: edgeR prefers integer-ish counts; numeric is okay but let's be explicit
storage.mode(counts) <- "numeric"

# quick check:
dim(counts)
#[1] 58674   393
head(rownames(counts), 10)
# [1] "5_8S_rRNA" "5S_rRNA"   "7SK"       "A1BG"      "A1BG-AS1"  "A1CF"      "A2M"       "A2M-AS1"   "A2ML1"    
# [10] "A2ML1-AS1"
head(colnames(counts), 10)
# [1] "53"  "59"  "88"  "91"  "133" "178" "190" "237" "238" "240"

```

5. Normalization
```{r}
# WHAT: Normalize raw counts with edgeR and compute logCPM
# WHY: Removes library size effects and provides comparable expression for scoring

library(edgeR)

dge <- DGEList(counts = counts)

# optional filtering (recommended)
keep <- filterByExpr(dge)
## Warning in filterByExpr.DGEList(dge) :
##  All samples appear to belong to the same group. 
#this warning is fine

dge <- dge[keep, , keep.lib.sizes = FALSE]

dge <- calcNormFactors(dge)
logCPM <- cpm(dge, log = TRUE, prior.count = 1)

dim(logCPM)
# [1] 16488   393
saveRDS(logCPM, file.path(res_dir, "logCPM_edgeR.rds"))

# load excel gene list
library(readxl)
gene_tbl <- read_excel(gene_xlsx) |> as.data.frame()
colnames(gene_tbl)
"Type I Genes"  "Type II Genes"
head(gene_tbl, 10)

```

(Debugging)
```{r}
# --- excel file debug and cleaning --- #
fix_excel_gene <- function(x) {
  x <- as.character(x)
  x <- gsub("\u00A0", " ", x)   # convert NBSP to normal space
  x <- gsub("[\r\n\t]", " ", x) # remove hidden breaks/tabs
  x <- trimws(x)               # trim leading/trailing
  x <- toupper(x)              # normalize case
  x
}

typeI_genes  <- unique(na.omit(fix_excel_gene(gene_tbl[["Type I Genes"]])))
typeII_genes <- unique(na.omit(fix_excel_gene(gene_tbl[["Type II Genes"]])))

# confirm the cleaning worked
typeI_genes[1:10]
"CD274" %in% typeI_genes

# check overlap
expr_genes <- toupper(rownames(logCPM))

typeI_in  <- intersect(typeI_genes, expr_genes)
typeII_in <- intersect(typeII_genes, expr_genes)

cat("Type I genes present:", length(typeI_in), "/", length(typeI_genes), "\n")
# Type I genes present: 139 / 144 
cat("Type II genes present:", length(typeII_in), "/", length(typeII_genes), "\n")
# Type II genes present: 42 / 50 
# ----------------------------------------------------------------------------------- #


# --- Recreate meta from the SOFT file --- #
soft_lines <- readLines(soft_file, warn = FALSE)

sample_start <- grep("^\\^SAMPLE", soft_lines)
sample_end   <- c(sample_start[-1] - 1, length(soft_lines))

parse_one_sample <- function(block) {
  gsm <- sub("^\\^SAMPLE = ", "", block[grep("^\\^SAMPLE", block)][1])

  title_line <- block[grep("^!Sample_title", block)]
  title <- if (length(title_line) > 0) sub("^!Sample_title = ", "", title_line[1]) else NA_character_

  chars <- block[grep("^!Sample_characteristics_ch1", block)]
  chars <- sub("^!Sample_characteristics_ch1 = ", "", chars)

  if (length(chars) == 0) return(tibble(sample = gsm, title = title))

  kv <- strsplit(chars, ":", fixed = TRUE)
  keys <- trimws(sapply(kv, `[`, 1))
  vals <- trimws(sapply(kv, function(x) paste(x[-1], collapse=":")))

  out <- as.list(vals); names(out) <- keys
  out$sample <- gsm
  out$title <- title
  as_tibble(out)
}

meta <- bind_rows(lapply(seq_along(sample_start), function(i) {
  block <- soft_lines[sample_start[i]:sample_end[i]]
  parse_one_sample(block)
}))

dim(meta)
head(meta[, 1:min(12, ncol(meta))])

# save 
saveRDS(meta, file.path(res_dir, "meta_from_SOFT.rds"))
write.csv(meta, file.path(res_dir, "meta_from_SOFT.csv"), row.names = FALSE)


# --- find the pSTAT3 field ---
grep("pstat3|stat3", colnames(meta), ignore.case = TRUE, value = TRUE)
# [1] "pstat3 status"

table(meta$`pstat3 status`, useNA = "ifany")
# negative positive 
#     156      237 

# --- Find which metadata column contains internal IDs (53, 59, 88, 91) ---
ids_to_find <- c("53", "59", "88", "91")

contains_id <- function(vec, id) {
  vec <- as.character(vec)
  any(grepl(paste0("(^|[^0-9])", id, "([^0-9]|$)"), vec, perl = TRUE))
}

cols_with_any_ids <- sapply(colnames(meta), function(cc) {
  vec <- meta[[cc]]
  any(sapply(ids_to_find, function(id) contains_id(vec, id)))
})

names(cols_with_any_ids)[cols_with_any_ids]
# [1] "title"

```

5. Run Singscore with delta = 0
```{r}
# Map UPPERCASE names back to the original rownames(logCPM)
map_upper_to_orig <- setNames(rownames(logCPM), toupper(rownames(logCPM)))

typeI_in_orig  <- unname(map_upper_to_orig[typeI_in])
typeII_in_orig <- unname(map_upper_to_orig[typeII_in])

# Score
ranked <- rankGenes(logCPM)

typeI_score  <- simpleScore(ranked, upSet = typeI_in_orig)$TotalScore
typeII_score <- simpleScore(ranked, upSet = typeII_in_orig)$TotalScore

delta <- typeI_score - typeII_score
Type_call <- ifelse(delta > 0, "Type I", "Type II")

calls <- data.frame(
  sample = colnames(logCPM),
  TypeI_score = typeI_score,
  TypeII_score = typeII_score,
  delta = delta,
  Type_call = Type_call,
  stringsAsFactors = FALSE
)

table(calls$Type_call)
# Type I Type II 
#    232     161 

# NOTE:
# 232 samples were called Type I
# 161 samples were called Type II
# 393 samples in total

# Save outputs
write.csv(calls, file.path(res_dir, "ALCL_TypeI_TypeII_calls.csv"), row.names = FALSE)
saveRDS(calls, file.path(res_dir, "ALCL_TypeI_TypeII_calls.rds"))
```


6. Compute Probabilistic Model
Note:
We have calls with TypeI_score, TypeII_score, delta, Type_call (from delta rule)
meta with pstat3 status and title = internal id
```{r}
# --- Merge + fit probabilistic model ---
library(dplyr)
library(stringr)

meta2 <- meta %>%
  mutate(internal_id = str_extract(title, "\\d+")) %>%
  mutate(internal_id = as.character(internal_id))

calls2 <- calls %>%
  left_join(meta2 %>% select(internal_id, `pstat3 status`, `alcl genetics`, `who classification`, tissue),
            by = c("sample" = "internal_id")) %>%
  mutate(truth_typeI = ifelse(`pstat3 status` == "positive", 1, 0))

# probabilistic model
# try 0.5
fit <- glm(truth_typeI ~ TypeI_score + TypeII_score, data = calls2, family = binomial())

calls2$prob_typeI <- predict(fit, newdata = calls2, type = "response")
calls2$Type_call_prob <- ifelse(calls2$prob_typeI >= 0.5, "Type I", "Type II")

table(calls2$Type_call_prob)
#  Type I Type II 
#     240     153
# That's number of samples classified as Type I vs Type II after the probabilistic model threshold (0.5)

# Check using confusion matrix & accuracy
tab <- table(Pred = calls2$Type_call_prob,
             Truth = ifelse(calls2$truth_typeI==1, "Type I", "Type II"))
tab
sum(diag(tab)) / sum(tab)
#          Truth
# Pred      Type I Type II
#  Type I     216      18
#  Type II     17     134

sum(diag(tab)) / sum(tab)
# [1] 0.9090909
```
Interpretation:
- Predicted pSTAT3 status from singscore scores with ~90.9% accuracy.
- On paper, prediction based on pSTAT3 was ~91% accurate.

Note regarding confusion matrix result:
216 samples: truth Type I AND predicted Type I 
134 samples: truth Type II AND predicted Type II 
18 samples: truth Type II but predicted Type I (false positives)
17 samples: truth Type I but predicted Type II (false negatives)
Accuracy = (216 + 134) / total = 350 / 385 = 0.909.

Why total is 385, not 393?
Because glm() automatically drops rows with missing values.
So likely 8 samples had NA in TypeI_score or TypeII_score or pstat3 status after merging (often from join issues or missing pstat3).

Simple check:
```{r}
sum(complete.cases(calls2[, c("TypeI_score","TypeII_score","truth_typeI")]))
# [1] 385
sum(is.na(calls2$`pstat3 status`))
# [1] 8

missing <- calls2 %>%
  filter(is.na(`pstat3 status`)) %>%
  select(sample, `alcl genetics`, `who classification`, tissue, Type_call_prob, prob_typeI)

missing
#   sample alcl genetics who classification tissue Type_call_prob prob_typeI
# 1    323          <NA>               <NA>   <NA>         Type I  0.9387034
# 2    371          <NA>               <NA>   <NA>         Type I  0.7163948
# 3    375          <NA>               <NA>   <NA>        Type II  0.1166157
# 4   2677          <NA>               <NA>   <NA>         Type I  0.9553168
# 5   3053          <NA>               <NA>   <NA>         Type I  0.7709503
# 6   3365          <NA>               <NA>   <NA>         Type I  0.9391672
# 7   5444          <NA>               <NA>   <NA>         Type I  0.5284725
# 8   7256          <NA>               <NA>   <NA>        Type II  0.1888479
```
Note:
- There are 393 samples total.
- 8 samples have pstat3 status = NA (missing in GEO metadata / SOFT parsing for those samples).
- So only 393 − 8 = 385 samples can be used as “truth-labeled” training/evaluation data.


(Debugging)
```{r}
expr_ids <- colnames(logCPM)
soft_ids <- meta2$internal_id

expr_ids <- expr_ids[!is.na(expr_ids)]
soft_ids <- soft_ids[!is.na(soft_ids)]

only_in_expr <- setdiff(expr_ids, soft_ids)
only_in_soft <- setdiff(soft_ids, expr_ids)

only_in_expr
# [1] "323"  "371"  "375"  "2677" "3053" "3365" "5444" "7256"
only_in_soft
# [1] "298"  "285"  "5197" "3540" "4444" "5031" "5619" "5131"


# Re-merge using only the matched samples
common_ids <- intersect(expr_ids, soft_ids)

calls_common <- calls %>% filter(sample %in% common_ids)

meta_common <- meta2 %>% filter(internal_id %in% common_ids)

calls2 <- calls_common %>%
  left_join(meta_common %>% select(internal_id, `pstat3 status`, `alcl genetics`, `who classification`, tissue),
            by = c("sample" = "internal_id")) %>%
  mutate(truth_typeI = ifelse(`pstat3 status` == "positive", 1, 0))

# Fit probabilistic model again (on clean matched set)
fit <- glm(truth_typeI ~ TypeI_score + TypeII_score, data = calls2, family = binomial())
calls2$prob_typeI <- predict(fit, newdata = calls2, type = "response")
calls2$Type_call_prob <- ifelse(calls2$prob_typeI >= 0.5, "Type I", "Type II")

tab <- table(Pred = calls2$Type_call_prob,
             Truth = ifelse(calls2$truth_typeI==1, "Type I", "Type II"))
tab
sum(diag(tab)) / sum(tab)
#          Truth
# Pred      Type I Type II
#   Type I     216      18
#   Type II     17     134
# [1] 0.9090909
```

8. Replicate figure B (on progress)
```{r}
# --- Build the column groups (ALK, BIA, TN-I, TN-II, DUSP22, TP63) ---
# Check WHO classification:
sort(unique(calls2$`who classification`))
# [1] "ALCLNEG" "ALCLPOS" "BIAALCL" "CUTALCL"

# Assign MolecularSubtype (paper panel blocks)
calls2 <- calls2 %>%
  mutate(
    MolecularSubtype = case_when(
      `who classification` == "BIAALCL" ~ "BIA",
      `alcl genetics` == "ALK" ~ "ALK",
      `alcl genetics` == "DUSP22" ~ "DUSP22",
      `alcl genetics` == "TP63" ~ "TP63",
      `alcl genetics` == "TN" & Type_call_prob == "Type I" ~ "TN-I",
      `alcl genetics` == "TN" & Type_call_prob == "Type II" ~ "TN-II",
      TRUE ~ "Other"
    )
  )

# keep only the ones in the figure blocks 
calls2_fig <- calls2 %>%
  filter(MolecularSubtype %in% c("ALK","BIA","TN-I","TN-II","DUSP22","TP63"))

table(calls2_fig$MolecularSubtype)
#   ALK    BIA DUSP22   TN-I  TN-II   TP63 
#   116     21     91     91     46     11 

# Note:
# Total: 376
# 376/385 matched samples fall into the Figure-B subtype blocks, and 9 samples are “Other” 
# (likely CUTALCL or something not included in that panel)


# --- Order the samples in the exact block order ---
subtype_order <- c("ALK","BIA","TN-I","TN-II","DUSP22","TP63")

calls2_fig <- calls2_fig %>%
  mutate(MolecularSubtype = factor(MolecularSubtype, levels = subtype_order)) %>%
  arrange(MolecularSubtype, desc(prob_typeI))   # nice separation inside each block

# --- Subset expression to those 376 samples, in that order ---
logCPM_fig <- logCPM[, calls2_fig$sample, drop = FALSE]
dim(logCPM_fig)
# [1] 16488   376
# 16488 genes = after edgeR filtering (filterByExpr)
# 376 samples = exactly the subtype blocks in Figure B (ALK/BIA/TN-I/TN-II/DUSP22/TP63)


# --- confirm gene list ---
library(readxl)
excel_sheets(gene_xlsx)
# [1] "Sheet1"
colnames(read_excel(gene_xlsx, sheet = 1))
# [1] "Type I Genes"  "Type II Genes"


# --- Add BIA + subtype labels ---
# check:
table(calls2_fig$MolecularSubtype)
#   ALK    BIA   TN-I  TN-II DUSP22   TP63 
#   116     21     91     46     91     11 

# check raw counts for the same samples, same order
counts_fig <- counts[, calls2_fig$sample, drop = FALSE]
dim(counts_fig)  
# [1] 58674   376


# --- Run edgeR supervised signatures ---
# NRC ≥ 50 reads in at least 0.7 × N cases (N = smaller group size)
# keep genes with |log2FC| ≥ 2 and FDR < 0.05
# direction = "up". Only keep genes up in the target subtype (makes “signature blocks” in the heatmap look like the paper)
get_subtype_signature <- function(counts_mat, subtype_vec, target_subtype,
                                  min_nrc = 50, frac = 0.7,
                                  lfc = 2, fdr = 0.05,
                                  direction = c("up","both")) {

  direction <- match.arg(direction)

  grp <- ifelse(subtype_vec == target_subtype, target_subtype, "Other")
  grp <- factor(grp, levels = c("Other", target_subtype))

  y <- DGEList(counts = counts_mat, group = grp)
  y <- calcNormFactors(y)

  # NRC filter (paper): NRC >= 50 in at least 0.7 * Nsmall samples
  n_other  <- sum(grp == "Other")
  n_target <- sum(grp == target_subtype)
  Nsmall <- min(n_other, n_target)
  req <- ceiling(frac * Nsmall)

  nrc <- cpm(y, normalized.lib.sizes = TRUE, log = FALSE)

  keep_gene <- rowSums(nrc[, grp == "Other",  drop = FALSE] >= min_nrc) >= req |
               rowSums(nrc[, grp == target_subtype, drop = FALSE] >= min_nrc) >= req

  y <- y[keep_gene, , keep.lib.sizes = FALSE]

  design <- model.matrix(~ grp)
  y <- estimateDisp(y, design)
  fit <- glmQLFit(y, design)
  qlf <- glmQLFTest(fit, coef = 2)

  tt <- topTags(qlf, n = Inf)$table
  tt <- data.frame(gene = rownames(tt), tt, check.names = FALSE)

  tt <- tt[tt$FDR < fdr & abs(tt$logFC) >= lfc, , drop = FALSE]
  if (direction == "up") tt <- tt[tt$logFC > 0, , drop = FALSE]

  tt
}

subtype_vec <- calls2_fig$MolecularSubtype
names(subtype_vec) <- calls2_fig$sample

alk_tt    <- get_subtype_signature(counts_fig, subtype_vec, "ALK",    direction="up")
bia_tt    <- get_subtype_signature(counts_fig, subtype_vec, "BIA",    direction="up")
dusp22_tt <- get_subtype_signature(counts_fig, subtype_vec, "DUSP22", direction="up")
tp63_tt   <- get_subtype_signature(counts_fig, subtype_vec, "TP63",   direction="up")

c(ALK=nrow(alk_tt), BIA=nrow(bia_tt), DUSP22=nrow(dusp22_tt), TP63=nrow(tp63_tt))
#    ALK    BIA DUSP22   TP63 
#      9     39     17     25 


# --- Freeze the subtype gene lists ---
topN <- function(tt, n=50) {
  tt <- tt[order(tt$FDR, -abs(tt$logFC)), , drop=FALSE]
  head(tt$gene, n)
}

ALK_genes    <- topN(alk_tt, 50)
ALK_genes #9
#[1] "ALK"    "IL1R1"  "GLIS3"  "IL1RAP" "CSMD2"  "TUBB3"  "PRF1"   "CEMIP"  "IL1R2"
BIA_genes    <- topN(bia_tt, 50)      
BIA_genes #39
# [1] "ITGAM"        "HSF4"         "CCL18"        "FN1"          "TIMP3"        "RNASE1"      
# [7] "RP1-309I22.2" "MXRA5"        "COL14A1"      "CCL4"         "F13A1"        "CCL3"        
# [13] "NLRP3"        "EGR2"         "CCL13"        "COL6A6"       "PTGIS"        "SLIT3"       
# [19] "MRC1"         "HMCN1"        "COL12A1"      "COMP"         "MME"          "COL22A1"     
# [25] "CD209"        "POSTN"        "FOS"          "C15orf48"     "FNDC1"        "COL11A1"     
# [31] "MARCO"        "FCGBP"        "SFRP4"        "THBS4"        "IL9"          "MAOA"        
# [37] "AGAP7P"       "COL6A5"       "RYR2"   
DUSP22_genes <- topN(dusp22_tt, 50)   
DUSP22_genes #17
# [1] "FBLN5"      "EZH2"       "NSMCE2"     "CNTNAP1"    "MSC"        "CCDC141"    "ZYG11A"    
# [8] "ALPK2"      "TTN"        "CCR8"       "MED12L"     "PHGDH"      "HNRNPA1P21" "BLK"       
# [15] "FERMT1"     "CCR4"       "NLGN4Y" 
TP63_genes   <- topN(tp63_tt, 50)     
TP63_genes #25
# [1] "TOB1"       "CCL5"       "ARSG"       "HGF"        "GNLY"       "TRGC2"      "IGHV1-69-2"
# [8] "TP63"       "GZMH"       "THEMIS"     "MYO7A"      "KIR3DL1"    "SNORD3B-2"  "IGHV3-30"  
# [15] "KIR2DL4"    "IGKV2-24"   "IGHV3-33"   "IGHV1-18"   "KIR3DL2"    "IGKV1-12"   "IGHV1-2"   
# [22] "IGLV2-34"   "AC233755.1" "IGLV3-19"   "PRG4" 

saveRDS(
  list(
    ALK = list(genes=ALK_genes, tt=alk_tt),
    BIA = list(genes=BIA_genes, tt=bia_tt),
    DUSP22 = list(genes=DUSP22_genes, tt=dusp22_tt),
    TP63 = list(genes=TP63_genes, tt=tp63_tt)
  ),
  file.path(res_dir, "paper_strict_subtype_signatures.rds")
)

# --- Build the full row blocks (Type I/II from Excel + subtype signatures) ---
gene_blocks <- list(
  "Type I"  = typeI_in_orig,
  "Type II" = typeII_in_orig,
  "ALK"     = ALK_genes,
  "BIA"     = BIA_genes,
  "DUSP22"  = DUSP22_genes,
  "TP63"    = TP63_genes
)

all_genes <- unique(unlist(gene_blocks))
all_genes <- intersect(all_genes, rownames(logCPM_fig))

mat <- logCPM_fig[all_genes, , drop = FALSE]

# z-score per gene (row) + clamp to match the paper’s color scale feel
zmat <- t(scale(t(mat)))
zmat[zmat > 10] <- 10
zmat[zmat < -10] <- -10

# -------------------------(debugging) ------------------------------------ #
# Build gene -> block mapping (first block wins if overlap)
gene2block <- setNames(rep(NA_character_, nrow(zmat)), rownames(zmat))

for (nm in names(gene_blocks)) {
  g <- intersect(gene_blocks[[nm]], rownames(zmat))
  gene2block[g[is.na(gene2block[g])]] <- nm
}

# Any genes still NA? (shouldn't happen, but just in case)
gene2block[is.na(gene2block)] <- "Other"

row_split <- factor(gene2block, levels = c(names(gene_blocks), "Other"))

# check
stopifnot(length(row_split) == nrow(zmat))
# -------------------------------------------------------------------------- #


# --- Plot Heatmap  ---
library(ComplexHeatmap)
library(circlize)
library(grid)

# 0) Required order (paper blocks)
subtype_order <- c("ALK","BIA","TN-I","TN-II","DUSP22","TP63")

# 1) Column split in the desired order
col_split <- factor(calls2_fig$MolecularSubtype, levels = subtype_order)

# 2) Make WHO a clean factor (keeps legend tidy)
who_levels <- c("ALCLNEG","ALCLPOS","BIAALCL","CUTALCL")
who_vec <- factor(calls2_fig$`who classification`, levels = who_levels)

# --- 3) Define WHO colors ---
who_cols <- c(
  ALCLNEG = "#00BFC4",  # teal
  ALCLPOS = "#2C7FB8",  # blue  (ALK+)
  BIAALCL = "#F0E442",  # yellow
  CUTALCL = "#F8766D"   # salmon
)

# --- 4) Define Subtype colors using your rules ---
subtype_cols <- c(
  ALK     = unname(who_cols["ALCLPOS"]),   # ALK same as ALK+
  BIA     = unname(who_cols["BIAALCL"]),   # BIA same as BIA
  `TN-I`  = "#CC79A7",
  `TN-II` = unname(who_cols["CUTALCL"]),   # TN-II same as Cutan.
  DUSP22  = "#7CAE00",
  TP63    = "#E69F00"
)


# --- 5) Rebuild top annotation WITH color mappings ---
ha <- HeatmapAnnotation(
  Subtype = col_split,
  WHO     = who_vec,
  col = list(
    Subtype = subtype_cols,
    WHO     = who_cols
  ),
  annotation_name_side = "right",
  annotation_legend_param = list(
    Subtype = list(title = "Molecular Subtype", at = subtype_order),
    WHO     = list(title = "WHO/ICC Subtype", at = who_levels)
  )
)

# 6) Color scale FIX (important!)
# My z quantiles showed most values are within about -2..+2,
# so I try saturate there to make the plot colors look like the paper.
col_fun <- colorRamp2(
  c(-10, -2, 0, 2, 10),
  c("green", "green", "black", "red", "red")
)

# 7) Plot
ht <- Heatmap(
  zmat,
  name = "Color Key",             
  col = col_fun,

  top_annotation = ha,

  row_split = row_split,
  column_split = col_split,

  cluster_rows = FALSE,
  cluster_columns = FALSE,

  show_row_names = FALSE,
  show_column_names = FALSE,

  # --- FIX row-split labels overlapping ---
  row_title_side = "left",
  row_title_rot = 90,                        # keep vertical
  row_title_gp = grid::gpar(fontsize = 11),  # smaller text helps
  row_gap = unit(3, "mm")                    # add space between blocks
)

png(file.path(fig_dir, "Figure6B_paper_reproduced.png"),
    width = 3200, height = 1800, res = 200)
draw(ht,
     heatmap_legend_side = "right",
     annotation_legend_side = "right",
     merge_legends = FALSE)
dev.off()

```



```{r}
sessionInfo()
```

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
